### 2021-10-22 documentos/vector_space.tex
Fixed everything.

\subsection{A vector space of channels} % {{{
\cpnote{Falta lo de las clases de equivalencias, pero aun no encuentro el
espacio. Quizá cuando haga falta se puede poner}
% \janote{Pienso que sí porque puede y debería ser útil para la sección siguiente
% de generadores.}
\par
% Inversion de la relacion entre tau y p, y que las probablidiades deben ser positivas {{{

@@

Let us start by recalling that the problem of determining complete positivity
of a PCE can be recast as determining which vectors $\vtau$ are mapped via
$\san$
% \begin{equation}
% \san\vec\tau=2^N\vec p
% \label{eq:condition}
% \end{equation}
to a vector with positive entries $\lambda_\valpha$, as in \eref{eq:lambda:is:A:tau}. 
% \janote{Siento que hace falta un enunciado por aquí, al inicio de la sección, que haga
% que el lector se pueda colgar de la idea de los espacios vectoriales. Quizás algo 
% como}\janote{We will show how to associate unequivocally PCE quantum
% channels with vector spaces.}
% . In components, this would read 
% \begin{equation}
% % \left(
% \sum_\vbeta
% \san_{\valpha\vbeta} \,
% % \right)
% \tau_{\vec \beta}=2^Np_{\vec\alpha}
% \label{eq:A:tau:is:p}
% \end{equation}
% where
% $\san_{\vec\alpha\vec\beta} =\sa_{\alpha_1\beta_1}\cdot\ldots\cdot
% \sa_{\alpha_N\beta_N}$. 
Using the fact that $a^{-1}=a/4$, so 
% (so $A = A^{-1}/2^N$) 
\begin{equation}
A^{-1} = \frac{1}{4^N} A,
\end{equation}
we can directly invert \eref{eq:lambda:is:A:tau} to obtain 
\begin{equation}
% \tau_{\vec \alpha}= \san_{\vec\alpha\vec\beta} \, p_\vbeta,
% \tau = 
% \san \lambda =2^N \tau
\sum_\vbeta \san_{\valpha \vbeta} \lambda_\vbeta =2^N \tau_\valpha
\label{eq:tau:is:A:p}
\end{equation}
which will serve as a starting point for our analysis. 
This is a remarkable equation, as it provides a method to diagonalize the
Choi matrix of any diagonal 
Pauli channel, as in \eref{eq:definition:Pauli}. 
% \janote{por qué diagonalizar al 
% canal de Pauli? Creo que sería más bien a la matriz de Choi, pero igual estoy 
% un poco pérdido con el sentido del enunciado. }

Other two simple but crucial
observation are the following. For valid quantum channels, 
% \begin{equation}
% % \sum_{
% \end{equation}
$\sum_{\valpha \in \Omega} \lambda_\valpha=0$ implies that 
$\lambda_\valpha=0$ for all $\valpha$ in an arbitrary subset of 
multi-indices $\Omega$, as each member of the sum is
greater or equal to zero, 
% $p_{\vec\alpha}+p_{\vec\beta}=0$ implies $p_{\vec\alpha}=p_{\vec\beta}=0$, as
% all $p_{\vec\alpha} \ge 0$
due to complete positivity of the underlying channel. 
Finally, looking into the first component of \eref{eq:tau:is:A:p}, and taking into 
account the normalization condition that $\tau_{\vec 0}=1$, we obtain
\begin{equation}
\sum_\valpha \lambda_\valpha = 2^N,
\label{eq:sum:alpha:is:2N}
\end{equation}
since, $A_{\vec 0, \vbeta}=1$ for all $\vbeta$. \par
% }}}
% Conexion de los indices para llegar al grupo de Klein {{{
Now we need a definition: to each multi-index $\vec\alpha$ we associate a \textit{set}
of multi-indices $\Phi(\vec\alpha)$ as follows
\begin{equation}
\Phi(\vec\alpha):=\left\{
\vec\beta:\san_{\vec\alpha\vec\beta}=1
\right\}
% =\left\{ \vec\beta: \sa_{\alpha_1\beta_1}\cdot\ldots\cdot \sa_{\alpha_N\beta_N}
% =1
% \right\}
\label{eq:8}
\end{equation}
% If we now assume that $\tau_{\vec\alpha}=1$, it follows from subtracting 
If we now assume that $\tau_{\vec\alpha}=1$, and calculate the difference between
\eref{eq:sum:alpha:is:2N} and
$\sum_{\vec\beta}\san_{\vec\alpha\vec\beta}\lambda_{\vec\beta}=2^N$
% \janote{la resta no es al revés?} 
(which follows from the aforementioned assumption), one obtains
% \begin{subequations}
% \begin{eqnarray}
% \sum_{\vec\beta}p_{\vec\beta}&=&1
% \label{eq:9a}\\
% \sum_{\vec\beta}\san_{\vec\alpha\vec\beta}p_{\vec\beta}&=&\tau_{\vec\alpha}=1
% \end{eqnarray}
% \label{eq:9}
% \end{subequations}
% that for all $\vec\beta\notin\Phi(\vec\alpha)$
\begin{eqnarray}
\lambda_{\vec\beta}=0,\quad \forall \vec\beta\notin\Phi(\vec\alpha).
\label{eq:condition:lambda:zero:Phi}
\end{eqnarray}
Thus, if $\tau_{\vec\alpha}=1$, then $\tau_{\vec\gamma}$
and $\tau_{\vec{\gamma^\prime}}$ are equal if
% , for all $\vec\beta\in\Phi(\vec\alpha)$.
\begin{equation}
\san_{\vbeta\vec\gamma}=\san_{\vbeta\vec{\gamma^\prime}}, \quad
\forall \vec\beta\in\Phi(\vec\alpha).
\label{eq:coneccion:dos:indices:A}
\end{equation}
% holds for all $\vec\beta\in\Phi(\vec\alpha)$. 
This follows from restricting the sum \eref{eq:tau:is:A:p}
to the $\lambda_\vbeta \ne 0$, given in \eref{eq:condition:lambda:zero:Phi}. 
% \janote{Para entender esto tuve que revisar de nuevo 
% el manuscrito de Francois. No veía cómo llegar a la implicación establecida
% en \eqref{eq:coneccion:dos:indices:A}. Creo que algunos pasos intermedios son
% importantes.}
Condition (\ref{eq:coneccion:dos:indices:A}) therefore connects three
multi-indices, $\valpha$,
$\vec\gamma$ and $\vec{\gamma^\prime}$.
When such a connection exists, $\tau_{\vec\alpha}=1$ implies
$\tau_{\vec\gamma}=\tau_{\vec{\gamma^\prime}}$. Let us now work out the nature
of this connection.

@@

It will be useful to think of the multi-index $\valpha$ as an element of 
a vector space. To do so, we notice that any group with the property that
$\alpha \oplus \alpha =0$ is indeed a vector space under the two-element field
$\{0,1\}$\cite{alguna:referencia}.  Then, we build the complete vector space,
with the same field, and defining $\valpha \oplus \vbeta = (\alpha_1 \oplus
\beta_1,\cdots, \alpha_N \oplus \beta_N)$
\footnote{
We can further \cpnote{desmechar} the vector space, by identifying each index
$\alpha \in \{0,1,2,3\}$ with its binary notation [e.g. identify $2$ with $(1,0)$]
such that to each multindex  $\valpha$ there corresponds a binary string of 
length $2N$. In fact, our sum $\oplus$ would correspond to addition modulo 2 of 
each of the $2N$ components. 
}. 
We can indeed restate \eref{eq:oplus}, an say that for valid quantum channels, 
if $\tau_{\vec\alpha}=1$, then
$\tau_{\vec\gamma}=\tau_{\vec\alpha\oplus{}\vec\gamma}$.
For example, in \fref{fig:two:qubit:examples}(c) the indices that 
correspond to preserved components are 
$\valpha^{(0)} = (0,0)$, 
$\valpha^{(1)} = (0,2)$, 
$\valpha^{(2)} = (1,0)$, 
$\valpha^{(3)} = (2,2)$, and
$\valpha^{(4)} = (3,2)$. However, 
$\valpha^{(1)} + \valpha^{(2)} = (1,2)$, which is not preserved and thus this 
diagram does not correspond to a quantum channel.  From this view 
we can derive several interesting observations that will be presented in 
the rest of the section.

@@

% Generacion de soluciones {{{
So we may now proceed to generate all solutions: we start out from the solution
having $\tau_{\vec0}=1$, with everything else $0$. We may then successively
switch $\tau$'s to $1$ for various values of $\valpha$, taking care
immediately to set equal to one the components of $\tau$ that correspond to values
of $\vbeta$ generated by the previously switched values of $\valpha$ via the
operation $\oplus$. Doing so, in an ordered way, allows one to general 
all valid quantum channels with a given set of preserved components, without the need
of exploring the exponentially large space of PCE.  \par 
% }}}}

@@

    \foreach \x in {0,1,2,3} {
      \foreach \y in {0,1,2,3} {
        \begin{scope}[shift={(\x,-\y)}] 
          \draw (0,0) rectangle (1,1); 
%           \node at (0.5,0.5) {$\tau_{\y,\x}$};
         \end{scope}
%         \node at (0,-\y) (input\y) {$i_\y$};
%         \node[block] at (2,-\y) (block\y) {$f_\y$};
%         \draw[->] (input\y) -- (block\y);
%         \draw[->] (block\y.east) -- +(0.5,0);
    }
    }

@@

% \begin{scope}[shift={(3,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(0,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(3,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}

@@

We can then see that all PCEs preserve $2^K$ components. In fact, with the above 
picture, it is clearly revealed that the vector space has dimension $2N$. Moreover, 

@@

elements of the vector space. This means that any PCE will preserve $2^K$ 

@@

\cpnote{Dejar que la raza elija uno de los dos parrafos anteriores} \par

 % }}}
% Cuantos canales conservan $2^K$ componentes? {{{
It is natural to ask how many PCEs exist that conserve $2^K$ components. One
can calculate such number, $\sss_{N,K}$, by examining the number of different
independent subsets of vectors that spawn a given vector subspace. In appendix 
\ref{app:contar} we show that 
\begin{equation}
\sss_{N,K}=\prod_{m=0}^{K-1} \frac{2^{2N-m}-1}{2^{K-m}-1}.
\label{eq:conteo:main}
\end{equation}
%An elementary test is $N=3$ and $K=2,3$: 
%\begin{eqnarray}
%\sss_{3,2}&=&\frac{(2^6-1)(2^5-1)}{(2^2-1)(2^1-1)}=\frac{63\cdot31}{3}=651,\\
%\sss_{3,3}&=&\frac{(2^6-1)(2^5-1)(2^4-1)}{(2^3-1)(2^2-1)(2^1-1)}=\frac{63\cdot31\cdot15}{7\cdot3}=1395,
%\end{eqnarray}
%which are indeed the values found numerically. 
From the above expression, it is easy to see the symmetry relation
\begin{equation}
\sss_{N,K}=\sss_{N,2N-K}
\end{equation}
which suggest an relation  between individual channels that for the time being
has escaped our efforts to identify. 
\par % }}}
% Frase de David {{{
\ddnote{Agrego esto:} A final remarks of the derivation as a tool. Concepts
might be complicated, but the tool gives some simple rules to the reader:
\textit{tell me what components you want to keep, and the (vector) summation
rules will give you the rest of components to make the target operation a CPTP
one.}
\par % }}}

>>>>

\subsection{A vector space of channels} % {{{
\cpnote{Falta lo de las clases de equivalencias, pero aun no encuentro el
espacio. Quizá cuando haga falta se puede poner}
\par
% Inversion de la relacion entre tau y p, y que las probablidiades deben ser positivas {{{

@@

Let us start by recalling that the problem of determining complete positivity
of a PCE can be recast as determining which vectors $\vtau$ are mapped via
$\san$
% \begin{equation}
% \san\vec\tau=2^N\vec p
% \label{eq:condition}
% \end{equation}
to a vector with positive entries $\lambda_\valpha$, as in \eref{eq:lambda:is:A:tau}. 
% . In components, this would read 
% \begin{equation}
% % \left(
% \sum_\vbeta
% \san_{\valpha\vbeta} \,
% % \right)
% \tau_{\vec \beta}=2^Np_{\vec\alpha}
% \label{eq:A:tau:is:p}
% \end{equation}
% where
% $\san_{\vec\alpha\vec\beta} =\sa_{\alpha_1\beta_1}\cdot\ldots\cdot
% \sa_{\alpha_N\beta_N}$. 
Using the fact that $a^{-1}=a/4$, so 
% (so $A = A^{-1}/2^N$) 
\begin{equation}
A^{-1} = \frac{1}{4^N} A,
\end{equation}
we can directly invert \eref{eq:lambda:is:A:tau} to obtain 
\begin{equation}
% \tau_{\vec \alpha}= \san_{\vec\alpha\vec\beta} \, p_\vbeta,
% \tau = 
% \san \lambda =2^N \tau
\sum_\vbeta \san_{\valpha \vbeta} \lambda_\vbeta =2^N \tau_\valpha
\label{eq:tau:is:A:p}
\end{equation}
which will serve as a starting point for our analysis. 
This is a remarkable equation, as it provides a method to diagonalize the
Choi matrix of any diagonal 
Pauli channel, as in \eref{eq:definition:Pauli}.

Other two simple but crucial
observation are the following. For valid quantum channels, 
% \begin{equation}
% % \sum_{
% \end{equation}
$\sum_{\valpha \in \Omega} \lambda_\valpha=0$ implies that 
$\lambda_\valpha=0$ for all $\valpha$ in an arbitrary subset of 
multi-indices $\Omega$, as each member of the sum is
greater or equal to zero, 
% $p_{\vec\alpha}+p_{\vec\beta}=0$ implies $p_{\vec\alpha}=p_{\vec\beta}=0$, as
% all $p_{\vec\alpha} \ge 0$
due to complete positivity of the underlying channel. 
Finally, looking into the first component of \eref{eq:tau:is:A:p}, and taking into 
account the normalization condition that $\tau_{\vec 0}=1$, we obtain
\begin{equation}
\sum_\valpha \lambda_\valpha = 2^N,
\label{eq:sum:alpha:is:2N}
\end{equation}
since, $A_{\vec 0, \vbeta}=1$ for all $\vbeta$. \par
% }}}
% Conexion de los indices para llegar al grupo de Klein {{{
Now we need a definition: to each multi-index $\vec\alpha$ we associate a \textit{set}
of multi-indices $\Phi(\vec\alpha)$ as follows
\begin{equation}
\Phi(\vec\alpha):=\left\{
\vec\beta:\san_{\vec\alpha\vec\beta}=1
\right\}
% =\left\{ \vec\beta: \sa_{\alpha_1\beta_1}\cdot\ldots\cdot \sa_{\alpha_N\beta_N}
% =1
% \right\}
\label{eq:8}
\end{equation}
% If we now assume that $\tau_{\vec\alpha}=1$, it follows from subtracting 
If we now assume that $\tau_{\vec\alpha}=1$, and calculate the difference between
\eref{eq:sum:alpha:is:2N} and
$\sum_{\vec\beta}\san_{\vec\alpha\vec\beta}\lambda_{\vec\beta}=2^N$
(which follows from the aforementioned assumption), one obtains
% \begin{subequations}
% \begin{eqnarray}
% \sum_{\vec\beta}p_{\vec\beta}&=&1
% \label{eq:9a}\\
% \sum_{\vec\beta}\san_{\vec\alpha\vec\beta}p_{\vec\beta}&=&\tau_{\vec\alpha}=1
% \end{eqnarray}
% \label{eq:9}
% \end{subequations}
% that for all $\vec\beta\notin\Phi(\vec\alpha)$
\begin{eqnarray}
\lambda_{\vec\beta}=0,\quad \forall \vec\beta\notin\Phi(\vec\alpha).
\label{eq:condition:lambda:zero:Phi}
\end{eqnarray}
Thus, if $\tau_{\vec\alpha}=1$, then $\tau_{\vec\gamma}$
and $\tau_{\vec{\gamma^\prime}}$ are equal if
% , for all $\vec\beta\in\Phi(\vec\alpha)$.
\begin{equation}
\san_{\vbeta\vec\gamma}=\san_{\vbeta\vec{\gamma^\prime}}, \quad
\forall \vec\beta\in\Phi(\vec\alpha).
\label{eq:coneccion:dos:indices:A}
\end{equation}
% holds for all $\vec\beta\in\Phi(\vec\alpha)$. 
This follows from restricting the sum \eref{eq:tau:is:A:p}
to the $\lambda_\vbeta \ne 0$, given in \eref{eq:condition:lambda:zero:Phi}. 
Condition (\ref{eq:coneccion:dos:indices:A}) therefore connects three
multi-indices, $\valpha$,
$\vec\gamma$ and $\vec{\gamma^\prime}$.
When such a connection exists, $\tau_{\vec\alpha}=1$ implies
$\tau_{\vec\gamma}=\tau_{\vec{\gamma^\prime}}$. Let us now work out the nature
of this connection.

@@

It will be useful to think of the multi-index $\valpha$ as an element of 
a vector space. To do so, we notice that any group with the property that
$\alpha \oplus \alpha =0$ is indeed a vector space under the two-element field
$\{0,1\}$\cite{alguna:referencia}.  Then, we build the complete vector space,
with the same field, and defining $\valpha \oplus \vbeta = (\alpha_1 \oplus
\beta_1,\cdots, \alpha_N \oplus \beta_N)$
\footnote{
We can further \cpnote{desmechar} the vector space, by identifying each index
$\alpha \in \{0,1,2,3\}$ with its binary notation [e.g. identify $2$ with $(1,0)$]
such that to each multindex  $\valpha$ there corresponds a binary string of 
length $2N$. In fact, our sum $\oplus$ would correspond to addition modulo 2 of 
each of the $2N$ components. 
}. 
We can indeed restate \eref{eq:oplus}, and say that for valid quantum channels, 
if $\tau_{\vec\alpha}=1$, then
$\tau_{\vec\gamma}=\tau_{\vec\alpha\oplus{}\vec\gamma}$.
For example, in \fref{fig:two:qubit:examples}(c) the indices that 
correspond to preserved components are 
$\valpha^{(0)} = (0,0)$, 
$\valpha^{(1)} = (0,2)$, 
$\valpha^{(2)} = (1,0)$, 
$\valpha^{(3)} = (2,2)$, and
$\valpha^{(4)} = (3,2)$. However, 
$\valpha^{(1)} + \valpha^{(2)} = (1,2)$, which is not preserved and thus this 
diagram does not correspond to a quantum channel.  From this view 
we can derive several interesting observations that will be presented in 
the rest of the section.

@@

% }}}
% Generacion de soluciones {{{
So we may now proceed to generate all solutions: we start out from the solution
having $\tau_{\vec0}=1$, with everything else $0$. We may then successively
switch $\tau$'s to $1$ for various values of $\valpha$, taking care
immediately to set equal to one the components of $\tau$ that correspond to values
of $\vbeta$ generated by the previously switched values of $\valpha$ via the
operation $\oplus$. Doing so, in an ordered way, allows one to generate
all valid quantum channels with a given set of preserved components, without the need
of exploring the exponentially large space of PCE.  \par 
% }}}}

@@

\cuadritotikz{}

@@

@@

We can then see that all PCE quantum channels preserve $2^K$ components. In
fact, with the above picture, it is clearly revealed that the vector space has
dimension $2N$. Moreover, 

@@

elements of the vector space. This means that any PCE quantum channel will preserve $2^K$ 

@@

\cpnote{Dejar que la raza elija uno de los dos parrafos anteriores}\janote{El
primero (simón, cambié de opinión)} \par

 % }}}
% Cuantos canales conservan $2^K$ componentes? {{{
It is natural to ask how many PCEs exist that conserve $2^K$ components. One
can calculate such number, $\sss_{N,K}$, by examining the number of different
independent subsets of vectors that spawn a given vector subspace. In appendix 
\ref{app:contar} we show that 
\begin{equation}
\sss_{N,K}=\prod_{m=0}^{K-1} \frac{2^{2N-m}-1}{2^{K-m}-1}.
\label{eq:conteo:main}
\end{equation}
%An elementary test is $N=3$ and $K=2,3$: 
%\begin{eqnarray}
%\sss_{3,2}&=&\frac{(2^6-1)(2^5-1)}{(2^2-1)(2^1-1)}=\frac{63\cdot31}{3}=651,\\
%\sss_{3,3}&=&\frac{(2^6-1)(2^5-1)(2^4-1)}{(2^3-1)(2^2-1)(2^1-1)}=\frac{63\cdot31\cdot15}{7\cdot3}=1395,
%\end{eqnarray}
%which are indeed the values found numerically. 
From the above expression, it is easy to see the symmetry relation
\begin{equation}\label{eq:conteo:simmetry}
\sss_{N,K}=\sss_{N,2N-K}
\end{equation}
which suggest a relation  between individual channels that
preserve $K$ and $2N-K$ Pauli components that for the time being
has escaped our efforts to identify. 
\par % }}}
% Frase de David {{{
\ddnote{Agrego esto:} A final remarks of the derivation as a tool. Concepts
might be complicated, but the tool gives some simple rules to the reader:
\textit{tell me what components you want to keep, and the (vector) summation
rules will give you the rest of components to make the target operation a CPTP
one.} \janote{Me gusta esto que puso David, pero lo movería a las conclusiones.}
\par % }}}
